A Computational Account of Default-Mode Function\\ by Control Theory and Reinforcement Learning

### 2016-09-25 paper/paper.tex
interm

\begin{abstract}
The default mode network (DMN) is the neural representation of the human
baseline mental activity.
%
Many research streams agree on its likely role in evolutionarily adaptive envisioning of past and hypothetical scenarios to predict the environment.
The DMN would hence be dedicated to continuous autobiographical memory retrieval, generate hypothetical outcomes,
and reward contingency evaluation when letting the mind go.
It explains its highest energy consumption in the brain and its intimate coupling with conscious awareness.
%
This concept paper proposes a process model that describes
\textit{how} the DMN may actually implement continuous
environmental assessment and prediction.
DMN function is recast from the mathematical perspective of control theory and
reinforcement learning.
Neurobiological evidence on the human and animal DMN
is linked to recent progress in autonomous systems research
in the machine-learning domain.
We argue that this engineering account can parsimoneously
explain a variety of existing neuroscientific hypotheses of DMN function.
%
Formal models of the computational processes subserved by the DMN
could offer statistical control of predictive human behavior.

@@

\section{The Neurobiological Properties of the Default-Mode Network}
mPFC and PCC = highest metabolic consumption in the brain/highest level of basal glucose energy consumption in humans

involvement of the PMC across the boundaries of classical cognitive domains

@@

\begin{itemize}
  \item
    vmPFC:
    \begin{itemize}
    \item
the vmPFC sub- serves predominantly non-ambiguous subjective-value-related evaluative processes
reward-informed and risk estimation of self-relevant environmental stimuli.
\item
the vmPFC might be more closely associated with orches- trating adapted behavior by bottom-up-driven processing of “what matters now”

\item
the vmPFC to monitoring others’ (Lotze et al., 2007) and one’s own (Lane et al., 1997; Phan et al., 2004) emotional responses, that is, other’s (external) emotional reac- tions and one’s own (visceral) arousal. Such real or imagined bodily states, believed to be represented in the vmPFC, proba- bly operate as a bioregulatory disposition governing cognition and decision making (Damasio, 1996; Nauta, 1971)

\item
Most importantly, independent whole-brain analyses from structural neuroimaging studies related the gray-matter volume (GMV) of the vmPFC to indices of social competence and social network complex- ity in both humans and monkeys (Lebreton et al., 2009; Powell et al., 2010; Lewis et al., 2011; Sallet et al., 2011). To our knowl- edge, none of these four correlations have been found yet for the dmPFC. Consequently, vmPFC, rather than dmPFC, anatomy appears to predict an individual’s social behavioral dispositions and social network properties,

\item
With respect to our seeds, inter-individual differences in social skills or social networks were most often related to morphological differences in the human and monkey vmPFC, in stark contrast to the dmPFC

\item
self-related behavior guided by stimulus evalua- tion and reward-learning, a voxel-based lesion-symptom map- ping (VLSM) study in 344 neurological patients demonstrated functional-anatomical specificity of the vmPFC for value-based decision-making (Gläscher et al., 2012).

\item
Glaescher: lesions -> value-based decision making  in 344 individuals
based decision-making, which included the orbitofrontal, ventro- medial, and frontopolar cortex.

\item
As to the vmPFC subnetwork, the GMV of the vmPFC and VS correlated with indices of social reward attitudes and behavior (Lebreton et al., 2009), concur- ring with vmPFC’s relation to the NAc and reward-related tasks.

\item
In studies of decision making and neuroeconomics, ventromedial prefrontal cortex (vmPFC) con- sistently reflects an individual’s subjective valuation (Behrens et al., 2008; Boorman et al., 2009; Knutson et al., 2005; Plass- mann et al., 2007)

\item
NAC was proposed to subserve model-free reinforcement learning in humans
(O'Doherty 2015)

\item
Schuck 2016: Performance within and across participants was related to state encoding in OFC;
“cognitive map” of task space in which the current state of the task is represented, and this representation is especially critical for behavior when states are unobservable from sensory input

\item
In both MACM and RS analy- ses, the vmPFC was more strongly connected with the nucleus accumbens (NAc), hippocampus (HC), posterior cingulate cor- tex (PCC), and retrosplenial cortex (RSC)
In both functional decoding analyses, the vmPFC was selectively associated with reward related tasks

\item
Indeed, the vmPFC, but not dmPFC, has been observed to have monosynap- tical connections with the ventral striatum (VS, which anatom- ically includes the NAc) in axonal tracing studies in monkeys (Haber et al., 1995; Ferry et al., 2000).
This is consistent with our results and probabilistic diffusion tensor imaging (DTI) trac- tography in humans and monkeys (Croxson et al., 2005) that quantified the VS to be substantially more likely connected to the vmPFC than dmPFC in both species.

\item
the vmPFC is preferentially connected with limbic and reward-related medial brain areas.
The NAc is thought to be linked to reward mechanisms that may not only modulate motivated behavior towards basic survival needs

\item
Such real or imagined bodily states, believed to be represented in the vmPFC, proba- bly operate as a bioregulatory disposition governing cognition and decision making (Damasio, 1996; Nauta, 1971),

\item
\end{itemize}

\item
  dmPFC:
  \begin{itemize}
    \item
dmPFC subserves predominantly ambiguous amodal metacogni- tive processes driven by top-down pathways

\item
sensory-independent de novo generation of meaning representations

\item
underlie inference, representation, and assessment
such as in one’s own and others’ mental states in functional neuroimaging research

\item
There is also activity in human dorsomedial prefrontal cortex that reflects expectations about what an observed person will do and errors in such predictions (Suzuki et al., 2012; van Schie et al., 2004)
largely sensory-independent, highly abstract (hence, less tangible) processes across time, space, and content domains.
hypothetical processing.

\item
A class of neurons in macaque dorsomedial frontal cortex, including the presupplementary area, responds preferentially to the actions that another macaque makes rather than their own actions (Yoshida et al., 2011)
Many of the same neurons differentiate between another macaque’s actions that were unsuccessful, because the circumstances changed, and actions that were unsuccessful because the other macaque did not make the best choice given the circum- stances. The macaques adjusted their own subsequent actions in different ways after these two types of obser- vations.

\item
 dmPFC (but not vmPFC) activity was related to the proficiency decline of mental state inference in elderly (Moran et al., 2012)

 \item
 congruently characterizes the dmPFC as a “mental sketchpad” (Goldman-Rakic, 1996) potentially impli- cated in modeling and binding plausible self- and other-related scenarios instructed by semantic concepts

 \item
Thus, mental state inference necessarily relies on the generation of probabilistic internal information. Supported by dmPFC’s functional association with episodic memory retrieval, such prima vista non-mnemonic construction processes are likely to be subserved by the neural network underlying retrieval of past and imagination of future scenes as indicated by recent neu- roimaging experiments and meta-analyses

scene con- struction processes when envisioning past, fictitious, and future events (Hassabis et al., 2007; Spreng et al., 2009; Bzdok et al., 2013).

\end{itemize}

  \item
    PMC (gradients):
    \begin{itemize}
      \item
      PMC myelinate comparatively late during postnatal development in monkeys (Goldman- Rakic, 1987). generally a sign of phylogenetic sophistication (Flechsig, 1920)
      \item
      (Patho-)Physiologically, meta- bolic fluctuations in the human PMC have been closely related to vari- ous instances of altered conscious awareness, including anesthesia (Fiset et al., 1999), sleep (Maquet, 2000), and restoration from vegeta- tive states (Laureys et al., 1999).      
      \item
      electrophys- iological research in animals implicated the PCC in strategic selection (Pearson et al., 2009), risk assessment (McCoy and Platt, 2005), and outcome-contingent behavioral modulation (Hayden et al., 2008), while the RSC was implicated in nav- igation and approach-avoidance behavior (Vann et al., 2009).

      \item
      Electrophysiologically, gamma band re- cordings in humans (Dastjerdi et al., 2011) and single-cell recordings in monkeys (Hayden et al., 2009) revealed
      \textit{activity reductions} in the PMC
      during task performance compared to rest.

      \item
      have long been speculated to reflect constant contemplation of (external) environment and (internal) memory, as an "inner minds eye"

      \item Bálint's syndrome: neurological disorder of conscious
      awareness resulting from damage in bilateral parietal cortex.
      inability to bind various individual features of the visual
      environment into an integrated whole (i.e., simultanagnosia)
      as well as to direct action towards 
      current unattended environmental objects
      (i.e., optic ataxia).

      This can viewed as an impairment in the gathering
      of information about alternative options (i.e., exploration) as well as
      leveraging these environmental opportunities
      (i.e., exploitation).

      \item RSC:
      \begin{itemize}
        \item
        Regarding memory processing, lesions in the RSC, as compared to other PMC regions, is most consistently associated with anterograde and retrograde memory impairments of different types of sensory infor- mation in rabbits (Gabriel and Talk, 2001) and humans (Rudge and Warrington, 1991; Valenstein et al., 1987).
        current findings and earlier lesion reports can be parsimoniously reconciled by the previously proposed notion that the RSC mediates between the organism's egocentric (i.e., focused on sensory input) and allocentric (i.e., focused on world knowledge) perspective frames (Burgess, 2008; Epstein, 2008; Valiquette and McNamara, 2007).

      \item
      neurons in posterior cingulate cortex signaled the distinction between trials on which monkeys pursued an exploratory rather than an exploitative strategy.

      \item
      Pearson2009:
      balance. Here we show that CGp neurons distinguish be- tween exploratory and exploitative decisions made by mon- keys in a dynamic foraging task.

      \item
      computation of environmental statistics (Pearson et al., 2009):

      single-cell recordings in the monkey PCC demonstrated this brain region's sensitivity to subjective target utility (McCoy and Platt, 2005) and integration across individual outcomes in decision making (Pearson et al., 2009).
      ng attraction or aversion to options with uncertain or risky rewards
      . CGp activation was better predicted by the
      subjective salience of a chosen target than by its actual value.

      -> internal currency of value

      motivated, targeted saccades hindering fixation of currently not
      attended features in the field of view (i.e., oculomotor apraxia).

      \item
      only cluster 2 was congruently coupled (across MACM and RSFC) with the amygdala (involved in significance evaluation) and the nucleus accumbens (involved in reward evaluation)
      s well as exclusively functionally asso- ciated with facial appraisal. Similarly, the human vPCC was the only PMC region connected to the laterobasal (rather than centromedial or superficial) nuclei group of the amygdala (Bzdok et al., 2013b), which is an amygdalar subregion probably devoted to continuously scanning environmental input for biological significance (Adolphs, 2010; Aggleton et al., 1980; Bzdok et al., 2011; Ghods-Sharifi et al., 2009; LeDoux, 2007).
      \item
      (McCoy and Platt, 2005):
      internal representations of subjective value lie at the very core of
      decision making,
      task where subjective value, but not objective value, varied, as indicated by
      subjects’ choices
      -> value matrix is not the same in each agent
      h the direction of impending
      Our data extend those
      findings, demonstrating a direct relationship between subjective preferences
      for uncertain rewards, or the opportunity to harvest a
      relatively large reward, and the activity of neurons thought to participate
      in the allocation of attention
      movement, as shown previously24,30, and the uncertainty of rewards
      associated with this movement
      CGp appears to carry spatial information that is scaled by
      subjective preferences for particular patterns of reward outcome.
      Our data indicate that CGp
      neurons signal subjective biases for uncertain rewards (or, perhaps, the
      potential to receive a large reward) rather than objective target value.
      the spatial sensitivity of CGp neurons was enhanced under
      conditions of high risk.
      -> delay discounting / non-immediate reward -> reinforcement learning

Such strategic decisions should incorporate not only recent reward history, but also opportunity costs and environmental statistics

\item
microstimulation in CGp promotes explo- ration of the previously antipreferred option.

\item
keys in a dynamic foraging task. Moreover, firing rates of these neurons predict in graded fashion the strategy most likely to be selected on upcoming trials. This encoding is distinct from switching between targets and is independent of the absolute magnitudes of rewards. These observations implicate CGp in the integration of individual outcomes across decision making and the modification of strategy in dynamic environments.
recorded the activity of single cingulate cortex (CGp) neurons

\item
outcome- contingent modulations in firing rate persisted into subsequent trials. Moreover, firing rate on any one trial predicted switching to the alternative option on the next trial. Finally, microstimulation in CGp follow- ing risky choices promoted a preference reversal for the safe option on the following trial. Collectively, these results demonstrate that CGp directly contrib- utes to the evaluative processes that support dynamic changes in decision making in volatile environments.

\item
IN SUM, these results suggest that the PCC is part of a network that monitors the subjective outcomes of possible decisions and integrates that information into higher-level strategies
**learn its current value and integrate this information with their statistical knowledge of the environment to predict its relative value on upcoming trials.**
\item
information processing in the CGp continuously adapts to changes in both the external environment and internal milieu.
\item
Hayden2008: adapativity of subjective reward schedules tht impact the action choices
lesion (Gabriel, 1990) studies strongly implicate CGp in working mem- ory, attention, and general arousal (Raichle and Mintun, 2006)— all processes that contribute to adaptive decision making. -> gradient defect
\item
integration of actions and their out- comes and thereby influences subsequent changes in behavior.
We confirmed a causal role for CGp in outcome evaluation and behavioral adjustment

\item HC:
\begin{itemize}
  \item
  Schachter/Buckner:
  The network described here is tied to the medial temporal lobe system, which is traditionally considered almost exclusively in the context of remembering the past. We suspect the adaptive value of
  episodic memory is not solely in its ability to afford mental reconstruction of the past but rather in its contribution to building mental models – simulations – of what might happen next or other perspectives on the immediate envir- onment, such as what others are thinking

  \item
The HC, in turn, is well known to be involved in memory and spatial navigation in animals and humans (von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).

\item
hippocampal damage [21,38–40] is associated with an impairment in thinking about one’s personal future

\item
bilateral hippocampal damage were markedly impaired relative to matched control subjects at imagining new fictitious experiences (Hassabis), their descriptions lacked spatial coherence and were instead fragmented and considerably less rich

\item 
A special case is the hippocampus whose highly recursive connection matrix is thought to function as a large ‘autoassocia- tor’, allowing the reconstruction of entire episodes from remembered fragments.

\item
 structural MRI studies of expert navigators (London taxi drivers) who show increased grey matter volume in posterior hippocampus seemingly at the expense of reduced grey matter volume in anterior hippocampus (Maguire et al. 2006b)

\item
hippocampal pyramidal cells during rest and sleep produce strongly coherent ensemble bursts believed to be critical in transferring information to the neocortex.

\item
statistical structure of hippocampal activity ->
spatial maps of surroundings are encoded
buszaki2004 theta oscillations:
Using large-scale recordings, researchers can follow complex patterns for extended time periods and determine whether their modification by experience will influence self-generated patterns in the absence of envi- ronmental inputs. Spike sequences, imposed upon the network by behavioral manipulations, recur spontaneously during subsequent sleep
indicating that neurons organize themselves into preferred cell assemblies, and the seeds of emergence stem from experi- ence-related activity.

Neurons, mostly studied in the hippocampus, which fire action potentials only when the animal is in a specific place in a given environment (that place where a cell fires is called the “place field” of that cell).

Numerous experiments have demonstrated that hippocampal neurons show place-related firing while the rat explores or traverses its environment so that each assembly of hippocampal principal cells defines a particular position of space

\item
Diba + Buzsaki 2007
We report that temporal spike sequences from hippocampal place neurons of rats on an elevated track recurred in reverse order at the end of a run, but in forward order in anticipation of the run, coinciding with sharp waves. Vector distances between the place fields were reflected in the temporal structure of these sequences. This bidirectional re-enactment of temporal sequences may contribute to the establishment of higher-order associations in episodic memory.

-> self-play in deepmind Go

\item
coding, transfer, and consolidation of information:
This emerging picture of hippocampal dy- namics suggests that neurons at the opposite ends of the distributions may convey differ- ent but complementary types of information. The ever-active minority of place cells may be responsible for generalizing across envi- ronments and affords the brain the capacity to regard no situation as completely un- known because every alley, mountain, river, or room has elements of previously experi- enced similar situations. In many situations,
this minority provides the “best guess” of the hippocampus and offers “good enough” solutions to get by. On the other hand, the majority of less active neurons constitute a large reservoir that can be mobilized (7) to precisely distinguish one situation from an- other and incorporate novel ones as distinct.

\item
damage to even the hippocampus alone is sufficient to cause amnesia +
envisoning themselves in future and even hypothetical scenarios/mentally constructing new experiences

The HC, in turn, is well known to be involved in memory and spatial navigation in animals and humans (von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).

\item
Hassabias2007:
Recollection of these rich autobiographical or episodic memories has been likened to mentally traveling back in time and re-experiencing one’s past (1). It has long been known that the hippocampus and related medial temporal lobe structures play a critical role in supporting episodic memory

In fact, episodic memory and imagining or constructing events share striking similarities in terms of the psychological processes engaged (19–21). These include imagery (22), sense of presence (1), retrieval of semantic information and multimodal details (23), and narrative structure (22).

visual imagery itself is not believed to be realized in the HC

\item
Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later

\end{itemize}

\item
the aRTPJ network is central for sensorimotor control by integrating supramodal stimulus-guided attention and action initia- tion during externally structured tasks.

\item
visuo- proprioceptive conflict (Balslev et al., 2005), and multi-modal detection of sensory changes (Downar et al., 2000)
brain areas connected to aRTPJ further responded to all changes of visual, auditory, or tactile stimulation in a multi-modal fMRI study (Downar et al., 2000).

\item
direct electrical stimulation of the RTPJ during neurosurgery was associated with altered perception and stimulus awareness (Blanke et al., 2002).

\item
the left and right TPJ further differ according to their cytoarchitectonic borders and gyral pattern (Caspers et al., 2008; Caspers et al., 2006)

\item
Neuroscientific investigations on hemispheric specialization have converged to the left cerebral hemisphere as dominant for language functions (Broca 1865, Wernicke 1874, Lichtheim 1885) and the right hemisphere as dominant for attention functions (Gazzaniga, Bogen et al. 1965, Sperry 1982, Stephan, Marshall et al. 2003). 
  \end{itemize}

  \item
    RTPJ:
    \begin{itemize}
\item
Sridharan: RTPJ involved in bottom-up saliency detection

\item
Corbetta2008:
Although both attentional networks have been most exten- sively investigated in vision, the available evidence indicates a supramodal function (Driver and Spence, 1998; Macaluso et al., 2002). The ventral network (right TPJ, right IFG) registers salient events in the environment not only in the visual but also in the auditory and tactile modalities (Downar et al., 2000)
Several lines of evidence indicate that two cortico-cortical neural systems are involved in attending to environmental stimuli (Corbetta and Shulman, 2002). A dorsal frontoparietal network. A second system, the ventral frontoparietal network, is not ac- tivated by expectations or task preparation but responds along with the dorsal network when behaviorally relevant objects (or targets) are detected (Corbetta et al., 2000). Both dorsal and ventral networks are also activated during reorienting, with en- hanced responses during the detection of targets that appear at unattended locations. For example, enhanced responses are observed when subjects are cued to expect a target at one location but it unexpectedly appears at another (i.e., ‘‘invalid’’ targets in the Posner spatial cueing paradigm)
or when a target appears infre- quently, as in ‘‘oddball’’ paradigms.

Intimately related to prediction error signaling:
An early theory of how the two networks interact (Corbetta and Shulman, 2002) proposed that when attention is reoriented to a new source 
of information (stimulus-driven reorienting), output from the ven- tral network interrupts (as a ‘‘circuit breaker’’) ongoing selection in the dorsal network, which in turn shifts attention toward the novel object of interest.

relevance is the critical factor that determines whether an object activates the ventral network (Downar et al., 2001). The ventral network might be considered a prime candi-date for mediating orienting to salient but unimportant stimuli, i.e., exogenous attention (Posner and Cohen, 1984), be- cause under passive conditions it is highly responsive to distinctive sensory
events in all modalities (Downar et al., 2000). The ventral network is well activated by stimuli that are important, even if they are not very distinctive. 
CURRENT OPINION: VA network does underlie  stimulus-driven reorienting to environmental stimuli based on their task relevance. An important conclusion from these neuroimaging studies is that the psychological dis- tinction between exogenous and endogenous orienting (Jo- nides, 1981) may not map onto different neural systems. "changing the control of behavior from one environ- mental input to another". Similar reorienting mechanisms may also be involved in shifting from a broad range of ‘‘internally directed’’ processes in order to deal with environmental events, as when interrupting memory retrieval (‘‘did I lock the car door?’’) to respond to a sudden stimulus (‘‘is that my cell phone ring- ing?’’). -> medites between the Default-Mode function and other more environmentally directed processing goals.
inducing network reconfiguration

\item
Bzdok2013

qualify RTPJ as a po- tential switch between exteroceptive and interoceptive mind sets implemented by their networks. The reciprocal nature of those mind sets is well illustrated by consistent reports of continuous shifting be- tween externally (i.e., sensory) and internally (i.e., self- and presumably social/memory-) oriented processing in controlled laboratory settings and daily routine (Smallwood et al., 2007).
\end{itemize}

\item
Corbetta2002:
 We suggest that the TPJ is also strongly engaged by stimuli that are behaviourally relevant but require a change in the current task set. 
 -> exogeneously induced state transition?
 <=> MCMC in HC / PCC -> slow? endogeneous changes in state transitions by the policy and value matrices
 For example, the alarm at the museum is not part of the task set of listening to the guide’s discussion of Hieronymous Bosch, but it is clearly a behaviourally relevant stimulus.

>>>>

\begin{abstract}
The default mode network (DMN) is the neural representation of the human
baseline mental activity.
%
Many research streams agree on its likely role in evolutionarily adaptive envisioning of past and hypothetical scenarios to predict the environment.
The DMN would hence be dedicated to continuous autobiographical memory retrieval,
generation of hypothetical outcomes,
and reward contingency evaluation when letting the mind go.
It explains its highest energy consumption in the brain and its intimate coupling with conscious awareness.
%
This concept paper proposes a process model that describes
\textit{how} the DMN may actually implement continuous
environmental assessment and prediction.
DMN function is recast from the mathematical perspective of control theory and
reinforcement learning.
Neurobiological evidence on the human and animal DMN
is linked to recent progress in autonomous systems research
in the machine-learning domain.
We argue that this engineering account can parsimoneously
explain a variety of existing neuroscientific hypotheses of DMN function.
%
Formal models of the computational processes subserved by the DMN
could offer statistical control of predictive human behavior.

@@

\section{Known Neurobiological Properties of the Default-Mode Network}
mPFC and PCC = highest metabolic consumption in the brain/highest level of basal glucose energy consumption in humans

scene con- struction processes when envisioning past, fictitious, and future events (Hassabis et al., 2007; Spreng et al., 2009; Bzdok et al., 2013).
it is not random but purposeful
draws on conceptual knowledge acquired through experience
operating on internal sources of information
Computational processes with a focus on the internal world,
when the environment is highly predictable and
external stimuli are irrelevant.
involvement of the PMC across the boundaries of classical cognitive domains

@@

The PMC myelinates relatively late during postnatal development in monkeys
(Goldman-Rakic, 1987), generally considered to
be a sign of phylogenetic sophistication (Flechsig, 1920).
Physiological and disturbed metabolic fluctuations in the
human PMC have been repeated related to
phenomena of changed conscious awareness,
including anesthesia (Fiset et al., 1999),
sleep (Maquet, 2000), and forms of coma (Laureys et al., 1999).      
%
The PMC has long been speculated to reflect constant computation of
the (external) environmental statistics and its (internal) representation
as an "inner minds eye" (Cavanna et al., 2006).
For instance, Bálint's syndrome is a neurological disorder of conscious
awareness that resulting from damage in the bilateral parietal cortex.
Patients are plaigued by an
inability to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as inability to direct action towards 
current unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in the gathering
of information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
The human PMC was congruently coupled in two functional connectivity modalities
with the amygdala
(involved in significance evaluation) and
the nucleus accumbens (involved in reward evaluation).
Specifically, the human ventral posterior cingulate cortex was
the PMC region most connected to the laterobasal
(rather than centromedial or superficial) nuclei group
of the amygdala (Bzdok et al., 2013b),
which is an amygdalar subregion probably devoted to
continuously scanning environmental input
for biological significance assessment
(Adolphs, 2010; Aggleton et al., 1980; Bzdok et al., 2011; Ghods-Sharifi et al., 2009; LeDoux, 2007).
%
Indeed,
electrophysiological recordings in animals implicated the PMC in
strategic selection (Pearson et al., 2009), risk assessment (McCoy and Platt, 2005), and outcome-contingent behavioral modulation (Hayden et al., 2008),
while its retrosplenial portion was
more specifically implicated in approach-avoidance behavior (Vann et al., 2009).
Neuron spiking activity in the PMC allowed distinguishing
whether a monkey would persue an exploratory or exploitative
behavioral strategy in a complex food foraging task (Pearson2009).
Single-cell recordings in the monkey PMC
demonstrated this brain region's sensitivity to
subjective target utility (McCoy and Platt, 2005) and integration
across individual decision-making instances (Pearson et al., 2009).
This DMN node encoded the
preference or aversion to options with uncertain reward outcomes
and its spiking activity was more associated with
subjectively perceived relevance of a chosen object
than by its factual value, based on an internal currency of value
In fact, direct stimulation of PMC neurons
promoted exploratory action towards options
with unsafe reward outcomes that were previously shunned.
Graded changes in firing rates of PMC neurons
that indicated changes in upcoming trials were
distinct from neuronal firing indicating switching between choices.
%
The retrosplenial portion of the PMC can subserve evaluation of reward
contingencies and action possibilities by integrating these with
information from memory and altered perspective frames.
Regarding memory retrieval, retrosplenial lesions have been
consistently associated with anterograde and retrograde memory impairments
of various kinds of sensory information
in rabbits (Gabriel and Talk, 2001) and humans
(Rudge and Warrington, 1991; Valenstein et al., 1987).
Regarding perspective frames, this PMC subregion has been
proposed to mediates between the organism's egocentric
(i.e., focused on sensory input) and
allocentric (i.e., focused on world knowledge) viewpoints
in animals and humans
(Burgess, 2008; Epstein, 2008; Valiquette and McNamara, 2007).
%
The PCC may consequently monitor the subjective outcomes
of possible decisions and integrates that information
with memory, perspective frames, and
reward schedules into higher-level strategies. 
Perceived value that differs across individuals updates
statistical knowledge of the environment
to predict delayed reward opportunities in the future.
In doing so, the PMC continuously adapts to changes
in both the external environment and internal milieu
that modulate strategic behavioral adjustment in volatile environments
to map possiblity spaces.

The dmPFC subserves predominantly ambiguous amodal processes
across time, space, and content domains in
sensory-independent top-down pathways.
This DMN node has been described as a “mental sketchpad” (Goldman-Rakic, 1996),
potentially implicated in de-novo generation and binding
of meaning representations instructed by stored semantics and memories.
%
The dmPFC may thus enable inference, representation, and assessment
of one's own and other individuals' action and thoughts.
For instance, neural activity in human dmPFC
reflected expectations about other peoples actions and errors thereof
(Suzuki et al., 2012; van Schie et al., 2004).
dmPFC activity indeed explained the proficiency decline
of inferring other peoples thoughts in the elderly (Moran et al., 2012).
Some dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, behavior
with fine-grained adjustment of contextual circumstances (Yoshida et al., 2011).
%
Such highly abstract neural computations necessarily rely on the
generation of probabilistic internal information drawing from
episodic memory retrieval, non-mnemonic scene construction processes,
and explicit knowledge of the external world.

The vmPFC subserves less ambiguous subjective-value-related evaluative processes
reward-informed and risk estimation of self-relevant environmental stimuli.
This DMN node is more closely associated with
orchestrating adapted behavior by bottom-up-driven
processing of “what matters now”.
Quantative lesion findings across  344 human individuals confirmed
a gross impairment in value-based decision making
(Glaescher et al., 2016).
The vmPFC is preferentially connected with limbic and reward-related areas.
The vmPFC has been observed to have monosynap- tical connections with the ventral striatum (VS, which anatom- ically includes the NAc) in axonal tracing studies in monkeys (Haber et al., 1995; Ferry et al., 2000).
The NAc is thought to be linked to reward mechanisms that may not only modulate motivated behavior towards basic survival needs, but also
subserve model-free reinforcement learning in humans more broadly
(O'Doherty 2015).
This is consistent with diffusion MRI (dMRI) tractography in humans and monkeys (Croxson et al., 2005) that quantified the VS to be substantially more likely connected to the vmPFC than dmPFC in both species.
Two functional connectivity modalities in humans strongly connected
the vmPFC with the nucleus accumbens (NAc), hippocampus (HC),
posterior cingulate cortex (PCC), and retrosplenial cortex (RSC).
%
The vmPFC is often proposed to be involved in (external) emotional
reactions and own (visceral) arousal.
Real or imaginged bodily states could be mapped in the vmPFC
as a bioregulatory disposition governing cognition
and decision making (Damasio, 1996; Nauta, 1971)
In studies of decision making and neuroeconomics, ventromedial prefrontal cortex (vmPFC) con- sistently reflects an individual’s subjective valuation (Behrens et al., 2008; Boorman et al., 2009; Knutson et al., 2005; Plass- mann et al., 2007)
This may be why performance within and across participants
was related to state encoding in OFC.
Such a “cognitive map” of action space was argued to encode
the current task state even when states are unobservable from sensory input,
which was shown to be critical for behavior.
The GMV of the vmPFC and VS correlated with indices of social reward attitudes and value-guided behavior (Lebreton et al., 2009).
Additionally,
independent whole-brain analyses from structural
neuroimaging studies related the gray-matter volume (GMV) of the vmPFC
(more consistently than any other
brain region) to indices of
social competence and social network complexity,
among the most complicated decision that humans and monkeys take
(Lebreton et al., 2009; Powell et al., 2010; Lewis et al., 2011; Sallet et al., 2011).
%

The HC is well known to be involved in memory and
spatial navigation in animals and humans
(von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).
Its highly recursive anatomical architecture
may be specifically designed to allowing reconstruction
whole episode of experience from memory fragments.
%
While the HC in the medial temporal lobe system
is traditionally believed to allow remembering the past,
there is now increasing evidence for a role
in constructing mental models in general (Schachter/Buckner).
Indeed,
hippocampal damage [21,38–40] is
not only associated with an impairment in reexperiencing the past (i.e., amnesia),
but also thinking about one’s own future and
imagining new fictitious experiences more broadly (Hassabis/PNAS).
Mental scenes created by hippocampus-lesioned patients exposed a lack of
spatial integrity, richness in detail, and overall coherence.
%
Single-cell recordings in the animal hippocampus revealed
some constantly active neuronal ensembles whose firing coincided with
distinct locations in space while the animal navigated through its surroundings.
London taxi drivers, individuals with high performance in spatial navigation,
were shown to exhibit increased grey matter volume in the
posterior hippocampus (Maguire et al. 2006b).
But encoding and generative reconstruction in the hippocampus extends
beyond mere spatial knowledge of the environment.
Based on large-scale recordings of hippocampal neuronal populations,
complex spiking patterns can be followed across extended periods including
their modification of input-free self-generated patterns
after environmental events (buszaki2004 theta oscillations).
Specific spiking sequences, that were elicited in experimental task conditions,
have been shown to be reenacted spontaneously during sleep.
Moreover, spike sequences measured in hippocampal place cells of rats
featured reoccurred directly after experimental trials
as well as directly before upcoming experimental trials (Diba + Buzsaki 2007).
Such hippocampal ensemble burts during rest and sleep
have been proposed to be critical in communicating local information
to the neocortex for long-term storage.
These mechanisms of the hippocampus probably make important contributions to the
recollection of autobiographical memory episodes and other
reexperienced or newly generated mental scenarios (Hassabias2007).
%
The HC thus orchestrates elements of experienced environmental aspects for
consolidations based on reenactment and for integration into 
rich mental scene construction. In this view, the HC my even influence
ongoing perception in the current environment (Maguire 2016).
-> self-play in deepmind Go
-> Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later

Finally,
the right and left TPJ are known to have hemispheric differences
according to their cytoarchitectonic borders and gyrification pattern
(Caspers et al., 2008; Caspers et al., 2006).
Neuroscientific investigations on hemispheric functional specialization
have converged to the right versus left cerebral hemisphere as dominant for
attention versus language functions (Broca 1865, Wernicke 1874, Lichtheim 1885;
Gazzaniga, Bogen et al. 1965, Sperry 1982, Stephan, Marshall et al. 2003).
%
The aRTPJ is a key node network is central for
action initiation during externally structured tasks and
sensorimotor control by integrating supramodal stimulus-guided attention
(Corbetta et al., 2002).
Involvement of this DMN node was repeatedly reported in
visuo-proprioceptive conflict (Balslev et al., 2005) and
multi-modal detection of sensory changes across
visual, auditory, or tactile stimulation in a multi-modal fMRI study
(Downar et al., 2000).
In humans, direct electrical stimulation of the
RTPJ during neurosurgery was associated with altered perception
and stimulus awareness (Blanke et al., 2002).
%
Importantly, the RTPJ has been shown to be intimately related to
supramodal prediction error signaling.
Neural activity in the RTPJ has been argued to be responsible 
for stimulus-driven attentional reallocation to 
salient and surprising sources of information
as a ‘‘circuit breaker’’ that recalibrates control and maintenance systems
(Bzdok et al., 2013; Corbetta et al., 2008).
In the face of large discrepancies between actual and previously predicted
environmental events the RTPJ acts a potential switch between
externally-oriented mind sets focussed on the
sensory world and internally-oriented mind sets focussed
on self-relevant mind-wandering of possible mental scenarios.
Transient RTPJ in humans for instance disrupted the
impact of predicted intentions of other individuals (Young et al. 2010b),
a capacity believed to be subserved by the DMN.
The RTPJ might hence be an important player that shifts away
from the default active ‘‘internally directed’’ processes
to deal with immediate environmental objects and contexts.
-> (slow) endogeneous changes in state transitions by the policy and value matrices
For example, the alarm at the museum is not part of the task set of listening to the guide’s discussion of Hieronymous Bosch, but it is clearly a behaviourally relevant stimulus. <=> MCMC in HC / PCC

The left TPJ in turn exhibits a close topographical relationship to
Wernicke's area
involved in the comprehension or understanding of written and spoken language.
Neurological patients with damage caused to Wernicke's area
have a major impairment of language comprehension
when listening to others or reading a book.
Their speech
preserves natural rhythm and about normal syntax, yet the
voiced sentences are devoid of meaning (i.e., aphasia).
Abstracting from the typical semantic interpretations in linguistics
and neuropsychology,
the LTPJ probably mediates access to and integrating of world knowlege.
For instance, LTPJ lesions also entail problems in recognizing
others' pantomimed action towards objects
without obvious relation to processing any language content
(Varney and Damasio 1987; Rothi et al. 1991; Buxbaum et al. 2005).
%
Also inner speech hinges on knowledge retrieval of statistical structure
about the physical and inter-personal world.
Multivariate volumetric analyses showed that the internal production of
formulated thought ("language of the mind") is closely related to the left TPJ
(Geva et al., 2011).
A number of neuroimaging studies demonstrated LTPJ activity responses
to sementically aberrant words in sentences, which suggests a role
in binding a given semantic concept into an integrated whole.
Further,
episodic memory recall and imagination strongly draws on
complex world knowledge.
Isolated building blocks of world statistics probably get reassembled
in internally generated visual scenarios that
navigate present action, weigh hypothetical possibilities, and forcast the future.
%
The LTPJ may hence facilitate the automated prediction of events
by incorporating experience-derived models of the world
into ongoing action, planning, and problem solving.
world -> policy matrix driven computation
self -> value matrix driven computation
